---
title: "ClinSight Configuration"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Deploy}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo = FALSE}
library(clinsight)
```

There two configuration files where the ClinSight configuration can be customized. [General configuration](#general-config) is stored in a configuration .yml file, while [study-specific configurations](#study-specific-config) is stored in a metadata Excel file. Both will be discussed in this vignette.

## General configuration {#general-config}

Below shows the current `config` file and the different configurations available:

```{r echo = FALSE, results ='markup', comment=""}
cat(readLines(system.file("golem-config.yml", package = "clinsight")), sep = "\n")
```

With the standard ('default') settings, internal example data with matching metadata will be used when running ClinSight.

When running with own test data, a different configuration needs to be set

### Shinymanager

### ShinyProxy deployment

### Posit Connect deployment

### Customizing configuration

If the profiles above are insufficient, you can also provide your own configuration file with fully customized settings with the steps described below.

To do so, first create a configuration template. You can use the following helper function to create

## Study-specific configuration {#study-specific-config}
