---
title: "Input Data Specification"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Input Data Specification}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
```{r setup}
library(clinsight)
```
In order to plug your organizations EDC data into the `clinsight` application,
you'll notice the `run_app()` function in `app.R` collects a `data` argument.
This vignette is all about the expected format of that `data` object / file.

Depending on the EDC vendor used, the size, shape, and format of their "raw
data" may vary. We've compiled a few functions that help admin users pre-process
the data they own. However, before we deep dive into how to use those, let's
focus on what the final result of those pre-processing steps in order to
understand what we need to start using the app.

## `Data` Specification

Baked into the `clinsight` package is an internal data set called
`clinsightful_data`. Here is a preview of that data:

```{r clinsightful_data}
# data("clinsightful_data") # Run to load the data into an R session. 
head(clinsight::clinsightful_data)
# colnames(clinsightful_data)
```

### Column specs

- `site_code`: 
- `subject_id`: 
- `event_repeat`: 
- `event_id`: 
- `event_name`: 
- `event_date`: 
- `form_id`: 
- `form_repeat`: 
- `edit_date_time`: 
- `db_update_time`: 
- `region`: 
- `day`: 
- `vis_day`: 
- `vis_num`: 
- `event_label`: 
- `item_name`: 
- `item_type`: 
- `item_group`: 
- `item_value`: 
- `item_unit`: 
- `lower_lim`: 
- `upper_lim`: 
- `significance`: 
- `reason_notdone`: 

For more information about 

## Pre-processing

### Raw Data

### Metadata

